A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-09-25 paper/paper.tex
minor fixes

The HC is well known to be involved in memory and
spatial navigation in animals and humans
(von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
Its highly recursive anatomical architecture
may be specifically designed to allowing reconstruction
whole episode of experience from memory fragments.
%
While the HC in the medial temporal lobe system
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general (Schachter/Buckner).
Indeed,
hippocampal damage [21,38–40] is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one’s own future and
imagining new fictitious experiences more broadly (Hassabis/PNAS).
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space while the animal navigated through its surroundings.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased grey matter volume in the
posterior hippocampus (Maguire et al. 2006b).
But encoding and generative reconstruction in the hippocampus extends
beyond mere spatial knowledge of the environment.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events (buszaki2004 theta oscillations).
Specific spiking sequences, that were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during sleep.
Moreover, spike sequences measured in hippocampal place cells of rats
featured reoccurred directly after experimental trials
as well as directly before upcoming experimental trials (Diba + Buzsaki 2007).
Such hippocampal ensemble burts during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage.
These mechanisms of the hippocampus probably make important contributions to the
recollection of autobiographical memory episodes and other
reexperienced or newly generated mental scenarios (Hassabias2007).
%
The HC thus orchestrates elements of experienced environmental aspects for
consolidations based on reenactment and for integration into 
rich mental scene construction. In this view, the HC my even influence
ongoing perception in the current environment (Maguire 2016).
-> self-play in deepmind Go
-> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later

Finally,
the right and left TPJ are known to have hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
(Caspers et al., 2008; Caspers et al., 2006).
Neuroscientific investigations on hemispheric functional specialization
have converged to the right versus left cerebral hemisphere as dominant for
attention versus language functions (Broca 1865, Wernicke 1874, Lichtheim 1885;
Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003).
%
The aRTPJ is a key node network is central for
action initiation during externally structured tasks and
sensorimotor control by integrating supramodal stimulus-guided attention
(Corbetta et al., 2002).
Involvement of this DMN node was repeatedly reported in
visuo-proprioceptive conflict (Balslev et al., 2005) and
multi-modal detection of sensory changes across
visual, auditory, or tactile stimulation in a multi-modal fMRI study
(Downar et al., 2000).
In humans, direct electrical stimulation of the
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness (Blanke et al., 2002).
%
Importantly, the RTPJ has been shown to be intimately related to
supramodal prediction error signaling.
Neural activity in the RTPJ has been argued to be responsible 
for stimulus-driven attentional reallocation to 
salient and surprising sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
(Bzdok et al., 2013; Corbetta et al., 2008).
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts a potential switch between
externally-oriented mind sets focussed on the
sensory world and internally-oriented mind sets focussed
on self-relevant mind-wandering of possible mental scenarios.
Transient RTPJ in humans for instance disrupted the
impact of predicted intentions of other individuals (Young et al. 2010b),
a capacity believed to be subserved by the DMN.
The RTPJ might hence be an important player that shifts away
from the default active ‘‘internally directed’’ processes
to deal with immediate environmental objects and contexts.
-> (slow) endogeneous changes in state transitions by the policy and value matrices
For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus. <=> MCMC in HC / PCC

>>>>

The HC is well known to be involved in memory and
spatial navigation in animals and humans
(von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).
Its highly recursive anatomical architecture
may be specifically designed to allowing reconstruction
whole episode of experience from memory fragments.
%
While the HC in the medial temporal lobe system
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general (Schachter/Buckner).
Indeed,
hippocampal damage [21,38–40] is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also thinking about one’s own future and
imagining new fictitious experiences more broadly (Hassabis/PNAS).
Mental scenes created by hippocampus-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal hippocampus revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space while the animal navigated through its surroundings.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased grey matter volume in the
posterior hippocampus (Maguire et al. 2006b).
But encoding and generative reconstruction in the hippocampus extends
beyond mere spatial knowledge of the environment.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events (buszaki2004 theta oscillations).
Specific spiking sequences, that were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during sleep.
Moreover, spike sequences measured in hippocampal place cells of rats
featured reoccurred directly after experimental trials
as well as directly before upcoming experimental trials (Diba + Buzsaki 2007).
Such hippocampal ensemble burts during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, which includes the DMN.
These mechanisms of the hippocampus probably make important contributions to the
recollection of autobiographical memory episodes and other
reexperienced or newly generated mental scenarios (Hassabias2007).
%
The HC thus orchestrates elements of experienced environmental aspects for
consolidations based on reenactment and for integration into 
rich mental scene construction. In this view, the HC my even influence
ongoing perception in the current environment (Maguire 2016).
-> self-play in deepmind Go
-> Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later

Finally,
the right and left TPJ are known to have hemispheric differences
according to their cytoarchitectonic borders and gyrification pattern
(Caspers et al., 2008; Caspers et al., 2006).
Neuroscientific investigations on hemispheric functional specialization
have converged to the right versus left cerebral hemisphere as dominant for
attention versus language functions (Broca 1865, Wernicke 1874, Lichtheim 1885;
Gazzaniga, Bogen et al. 1965, Sperry 1982, Stephan, Marshall et al. 2003).
%
The aRTPJ is a key node network is central for
action initiation during externally structured tasks and
sensorimotor control by integrating supramodal stimulus-guided attention
(Corbetta et al., 2002).
Involvement of this DMN node was repeatedly reported in
visuo-proprioceptive conflict (Balslev et al., 2005) and
multi-modal detection of sensory changes across
visual, auditory, or tactile stimulation in a multi-modal fMRI study
(Downar et al., 2000).
In humans, direct electrical stimulation of the
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness (Blanke et al., 2002).
%
Importantly, the RTPJ has been shown to be intimately related to
supramodal prediction error signaling.
Neural activity in the RTPJ has been argued to be responsible 
for stimulus-driven attentional reallocation to 
salient and surprising sources of information
as a ‘‘circuit breaker’’ that recalibrates control and maintenance systems
(Bzdok et al., 2013; Corbetta et al., 2008).
In the face of large discrepancies between actual and previously predicted
environmental events the RTPJ acts a potential switch between
externally-oriented mind sets focussed on the
sensory world and internally-oriented mind sets focussed
on self-relevant mind-wandering of possible mental scenarios.
Transient RTPJ in humans for instance disrupted the
impact of predicted intentions of other individuals (Young et al. 2010b),
a capacity believed to be subserved by the DMN.
The RTPJ might hence be an important player that shifts away
from the default active ‘‘internally directed’’ processes
to deal instead with immediate environmental objects and contexts.
-> (slow) endogeneous changes in state transitions by the policy and value matrices
For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus. <=> MCMC in HC / PCC
